---
title: "R Notebook"
output: html_notebook
---

2.	Body measurement

Adapt some code from the Physio_dash.R shiny app to create a function and code that allows you to create some breathing rate simulated data. Then adapt some more code again  from the Physio_dash.R shiny app, using the dygraphs and xts packages, to create a script that plots a dygraph plot and print the dygraph plot of the simulated data.

Write the simulated data to a datafile so that it can be used to check the graph.

Submit an R script with the code you used, a pdf of the plot and the datafile of your simulated data.

#Loading required libraries
```{r}
library(tidyverse)
library(dygraphs)
library(xts)
```

#ADAPTED CODE:
  data_HR_sim <- eventReactive(input$simulate_data,{
    time <- seq(t_start, t_start+(input$duration*60000), by = 500)
    Heartrate <- as.numeric(arima.sim(model = list(ar = input$autocorrelation) , n = length(time), mean = 8, sd = 5))
    data.frame(time,Heartrate)

BR by = 1500, mean = 4, sd = 5, variable name = Breathing.Rate
#Creating a function to simulate data

The function has parameters of  t_start, duration and autocorrelation. The time variable generates the time from the the starting point(t_start) to the present time. 
```{r}
simulate_ft<- function(
    t_start,
    duration,
    autocorrelation
){
  #creates a vector that runs from 0  with 1500 increments. This is the time sequence for our simulation data.
  time <- seq(t_start, t_start+(duration*60000), by = 1500)
  
  Breathing_Rate <- as.numeric(arima.sim(model = list(ar =autocorrelation) , n = length(time), mean = 4, sd = 5))
#The  value for the Breathing_Rate is obtained using the arima(AutoRegressive Integrated Moving Average) function that takes in a input of AR and MA coefficients as a list [model = list(ar = autocorrelation)] and we also define the length of the output series as the length of the time generated(n = length(time)) with a mean of 4 and standard deviation of 5(-Physio_dash.R )
  
  Breathing_Rate<- abs(Breathing_Rate) #As there should be no negative values in breathing, taking the absolute value
  Breathing_Rate<- data.frame(time,Breathing_Rate) #creates a data frame of time and Breathing_Rate as the output
  return(Breathing_Rate) #Return resulting data as output of function
}
```

#Simulating data using the function
The values for the t_start,duration and autocorrelation are passed to the function to generate the Breathing_Rate

t_start=0
duration=4
autocorrelation=0.8
```{r}

sim_data<-simulate_ft(0,4,0.8)
sim_data #Creating
```

#ADAPTING CODE FOR PLOTTING
  data_HR <- reactive({
    if (!is.null(data_HR_sim())){
      data <- data_HR_sim()
    }  
    data <- data %>%
      dplyr::select(time,Heartrate) %>%
      dplyr::mutate(biometric = "HR") %>%
      dplyr::rename(value = Heartrate) %>%
      dplyr::mutate(time_date = as.POSIXct(as.numeric(as.character(time))/1000, origin = "1970-01-01", tz="Europe/London"))
    return(data)
  })
  
  output$plot_HR <- renderDygraph({
    HR_xts <- xts(x = data_HR()$value, order.by = data_HR()$time_date) # specific dataframe with time series class
    dygraph(HR_xts, main = " ", ylab = "Heart Rate (bpm)", xlab = "time", group = "biometrics")%>%
        dySeries("V1", label = "Heart Rate (bpm)") %>%
        dyRangeSelector(height = 20) %>%
        dyOptions(colors = 'red') %>%
        dyUnzoom() %>%
        dyLegend(width = 300)
  })
```{r}
plot_data <- sim_data %>%
      dplyr::select(time,Breathing_Rate) %>%
      dplyr::mutate(biometric = "BR") %>%
      dplyr::rename(value = Breathing_Rate) %>%
      dplyr::mutate(time_date = as.POSIXct(as.numeric(as.character(time))/1000, origin = "1970-01-01", tz="Europe/London"))
plot_data

#WRANGLING DATA. 
#Data wrangling is done using the dplyr package . It takes the data provided by plot_data and changes it to create a data frame with four variables time, value, biometric, and time_date. It retains the original numeric time data as time, changes the breathing rate variable to value and gives it a label of BR in another variable. Finally the last line of wrangling code creates a POSIX compatible time variable using the mutate command.
```

#Writing the data into a csv file

```{r}
write.csv(plot_data,"/Users/martinchristy/Desktop/BA model prac/Simulated_Data.csv")
```

#Plotting 
```{r}
BR_xts <- xts(x = plot_data$value, order.by = plot_data$time_date) # specific dataframe with time series class
#
 dygraph(plot_data, main = " ", ylab = "Breathing Rate (rpm)", xlab = "time", group = "biometrics")%>%
   dySeries("value", label = "Breathing Rate (rpm)") %>%
   dyRangeSelector(height = 20) %>%
   dyOptions(colors = 'BLUE') %>%     
   dyUnzoom() %>%     
   dyLegend(width = 300)    
```



