---
title: "R Notebook"
output: 
  html_notebook: 
    fig_caption: yes
---

2.	Body measurement

Adapt some code from the Physio_dash.R shiny app to create a function and code that allows you to create some breathing rate simulated data. Then adapt some more code again  from the Physio_dash.R shiny app, using the dygraphs and xts packages, to create a script that plots a dygraph plot and print the dygraph plot of the simulated data.

Write the simulated data to a datafile so that it can be used to check the graph.

Submit an R script with the code you used, a pdf of the plot and the datafile of your simulated data.

```{r}
library(dygraphs)
library(tidyverse)
library(dplyr)
library(xts)
```

#ADAPTED CODE:

  data_HR_sim <- eventReactive(input$simulate_data,{
    time <- seq(t_start, t_start+(input$duration*60000), by = 500)
    Heartrate <- as.numeric(arima.sim(model = list(ar = input$autocorrelation) , n = length(time), mean = 8, sd = 5))
    data.frame(time,Heartrate)
  })
  

    
    
#Creating a function
```{r}
sim_data <- function(
  t_start,
  duration,
  autocorrelation
){
  time <- seq(t_start, t_start+(duration*60000), by = 1500)
  Breathing_Rate <- as.numeric(arima.sim(model = list(ar = autocorrelation) , n = length(time), mean = 4, sd = 5))
  Breathing_Rate<- abs(Breathing_Rate)
  Breathing_Rate<- data.frame(time,Breathing_Rate)
  return(Breathing_Rate)
}
```


Simulate some data
```{r}
df<-sim_data(1500000000000,1,0.5)
df

```
 
#Writing the data as a csv file  
```{r}
write.csv(df,"C:/Users/Jidhin/Desktop/Q2_breathing_sim_data.csv")

```

#ADAPTED CODE: 
  output$plot_HR <- renderDygraph({
    HR_xts <- xts(x = data_HR()$value, order.by = data_HR()$time_date) # specific dataframe with time series class
    dygraph(HR_xts, main = " ", ylab = "Heart Rate (bpm)", xlab = "time", group = "biometrics")%>%
        dySeries("V1", label = "Heart Rate (bpm)") %>%
        dyRangeSelector(height = 20) %>%
        dyOptions(colors = 'red') %>%
        dyUnzoom() %>%
        dyLegend(width = 300)
  })
  
```{r}
data <- df
data <- data %>%
  dplyr::select(time,Breathing_Rate) %>%
  dplyr::mutate(biometric = "BR") %>%
  dplyr::rename(value = Breathing_Rate) %>%
  dplyr::mutate(time_date = as.POSIXct(as.numeric(as.character(time))/1000, origin = "1970-01-01", tz="Europe/London"))
data
```
  
#Plot#

HR_xts <- xts(x = data_HR()$value, order.by = data_HR()$time_date) # specific dataframe with time series class
    dygraph(HR_xts, main = " ", ylab = "Heart Rate (bpm)", xlab = "time", group = "biometrics")%>%
        dySeries("V1", label = "Heart Rate (bpm)") %>%
        dyRangeSelector(height = 20) %>%
        dyOptions(colors = 'red') %>%
        dyUnzoom() %>%
        dyLegend(width = 300)
        
```{r}
BR_xts <- xts(x = data$value, order.by = data$time_date)
 dygraph(data, main = " ", ylab = "Breathing Rate (bpm)", xlab = "time", group = "biometrics")%>%
        dySeries("value", label = "Breathing Rate (rpm)") %>%
        dyRangeSelector(height = 20) %>%
        dyOptions(colors = 'BLUE') %>%
        dyUnzoom() %>%
        dyLegend(width = 300)



```

   
    